### 2020-01-27 paper4/article.tex
Paper 4 added.

%\documentclass[12pt, twocolumn]{article}
%\usepackage{helvet}
%
%\usepackage{epsfig}
%\usepackage[latin1]{inputenc}
%\begin{document}
%
%\title{Emulating Von Neumann Machines and Massive Multiplayer Online Role-
%Playing Games}
%\author{Mickey Mouse, Goofy G. Goof and Donald Duck}
%
%\date{}

%\maketitle

%\section*{Abstract}
%
% Many computational biologists would agree that, had it not been for
% Byzantine fault tolerance, the synthesis of replication that made
% developing and possibly investigating erasure coding a reality might
% never have occurred. In this work, we prove  the synthesis of linked
% lists. Even though such a hypothesis at first glance seems
% counterintuitive, it always conflicts with the need to provide
% object-oriented languages to systems engineers. APER, our new framework
% for mobile archetypes, is the solution to all of these grand
% challenges.

In \cite{Batchelor} we believe.

%\begin{footnotesize}
%\bibliography{scigenbibfile.Donald+Duck.Mickey+Mouse.Goofy+G.+Goof}\bibliographystyle{acm}
%\end{footnotesize}
%
%\end{document}

>>>>

\section{Introduction}

Small particles, droplets and bubbles are ubiquitously present in flowing fluids. When a suspended particle is transported by a viscous fluid, it modifies the flow around it. If another particle happens to be in the region of the modified flow, mutual hydrodynamic interactions between the particles will take place. The interactions are given implicitly by imposing boundary conditions on the flow that must hold simultaneously on the surfaces of all interacting particles \citep{hb,kim_karrila}. This setting is inconvenient for analyses, both theoretical and numerical. Thus there is no answer to even simplest questions, for instance whether there can be a non-trivial stationary configuration of particles that would flow as a whole due to the hydrodynamic interaction. Although these interactions somewhat resemble electrostatic interactions, there is no a hydrodynamic counterpart of Earnshaw's theorem \citep{pur} stating that such simple configurations are impossible. Here we provide an example of this possibility in the presence of a boundary and demonstrate that boundaries can have surprising and  non-trivial effects on hydrodynamic interactions.

The only well-studied case of hydrodynamic interactions of particles transported by non-uniform flow is the case of two particles in a time-independent low Reynolds number linear flow. This was studied in the seminal \cite{batchelor_green_1972} paper, see also \cite{arp} for more details and an account of various contributions to the problem. If the particles' (and fluid) inertia can be entirely neglected, the vector between the particle centers obeys an autonomous first-order evolution equation, which gives its rate of change as a function of the instantaneous value. The use of the symmetries makes it possible to quantify the interaction by the two scalar functions of the distance, which have been tabulated \citep{batchelor_green_1972}, see also \cite{kim_karrila}. This case presents no stationary configurations for the two particles. One of the main applications is the Poiseuille flow in the channel shown in Fig. \ref{fig:setup}. If both particles are far from the walls, $z_0 \gg a, r$, they can seemingly be considered as flowing in an unbounded shear flow and the analysis of \cite{batchelor_green_1972,arp,lin_lee_sather_1970} applies. The BG theory thus predicts that there are no possible stationary configurations of the particle pair. Here we demonstrate that the approximation of an unbounded flow overlooks such configurations and also other phenomena, which hold independently of how large $z_0$ is, cf. \cite{zurita-gotor_2007}. Thus, the presence of the wall exhibits a singular perturbation of the BG theory.

Recently, stationary configurations of particles transported in microfluidic channels attracted attention due to the possibility of flow-assisted microfabrication by using a combination of hydrodynamic and non-hydrodynamic (\ie adhesive) interactions \citep{tab0,tabeling}. Under certain conditions, the particles can self-assemble into clusters of different morphology that flow with no change of inter-particle distances, see \cite{flow-assist} for detailed discussions. These micron-scale clusters can then be solidified and collected from the flow, and be potentially used for fabrication of functional metamaterials. To theoretically explain and subsequently predict the structure formation of the suspended particles observed in experiments, \cite{tabeling} proposed a
model based on a dipolar asymptotic form of hydrodynamic interaction. Notice that the dipolar form only holds at large particle separations (it was derived in detail using the fundamental solution for the channel flow \citep{LironMochon}, see \cite{Fouxon_2017} and also references therein). One reason for the apparent applicability of this description, despite the particles were \textit{close to} each other, is likely the dominance of the adhesive radial forces between particles at close proximity, such that any hydrodynamic interaction producing a non-zero tangential velocity component would yield a similar cluster formation dynamics (see \cite{flow-assist} for further evidence). In contrast, a consistent predictive theory of hydrodynamic interactions should hold irrespective of the presence of adhesive forces, and allow for analysis of interaction of flowing particles at close proximity and near the wall, as in experimental setup \citep{tabeling}. The present paper is a step toward this theory.
%
\begin{figure}[t]
 \centering
 \includegraphics[width=0.75\textwidth]{setup.png}
 \caption{Setup of a particle pair in the Poiseuille flow (depicted in a comoving reference frame). In this work we study the case $z_0 \gg a$, where $a$ is the particle radius, however does not necessarily require $z_0\gg r$. The upper wall is assumed to be much further away from the particle pair than $z_0$.
%The BG theory describes well the evolution when $z_0\gg r$ however, for open trajectories predicts that the particles separate...
}
 \label{fig:setup}
\end{figure}

There are two differences between the channel flow and the unbounded shear flow considered in the BG theory  \citep{lin_lee_sather_1970,batchelor_green_1972}. The velocity profile of the channel flow is quadratic \citep{quadratic} in the coordinate rather than linear. This difference is often irrelevant when the interacting particles are located much closer to one of the channel walls, so that the flow can be closely approximated by the linear shear flow. This is the case we consider in the present paper. Another difference, is that the no-slip rigid wall is always at a finite distance and it interacts with the flowing particles.

We first consider the evolution of the inter-particle distance when the effect of the wall is neglected and the BG theory applies. It is useful to consider the three-dimensional phase space spanned by all possible distances $\bm r$ between the spheres' centers where one of them is at the origin. The distance $\bm r(t)$ between the spheres' centers obeys an autonomous evolution equation which means that there is a well-defined phase space flow $\bm V^0(\bm r)$ such that $\dot{\bm r}=\bm V^0(\bm r(t))$ and a unique trajectory passes through each point. This is the consequence of neglecting particles' and fluid inertia and the translational invariance due to which the shear resistance matrix depends on $\bm r$ only, cf. \cite{sh} (translation in a linear flow changes the flow by a constant vector, irrelevant by Galilean invariance). The flow $\bm V^0(\bm r)$ does not vanish anywhere so that there are no steady configurations. The absence of critical points with $\bm V^0(\bm r)=0$  implies a simple structure of the phase space. This can be most readily observed in the symmetry plane formed by the horizontal flow direction $x$ and the vertical velocity gradient direction $z$ (see Fig.~\ref{fig:SeparatrixBG}). The trajectories that belong to the plane never leave it, $V_y(y=0)=0$, and can be considered separately. There is a simple dichotomy of the trajectories:  closed trajectories crossing the $x$-axis and open trajectories that do not cross the $x$-axis. The open trajectories describe the faster particle overtaking the slower one. The particles return to their original vertical positions following the hydrodynamic encounter and there is fore-and-aft symmetry of the phase portrait. In contrast, the trajectories that cross the $x$-axis are closed, corresponding to a bound pair of spheres orbiting around each other. Open and closed trajectories are separated by the separatrix that touches the $x$-axis asymptotically at large distances \citep{lin_lee_sather_1970}. Rotation of this separatrix around the $z$-axis creates an axisymmetric surface that separates the regions of open and closed trajectories in space (it is not readily evident how this axial symmetry could be guessed \textit{a priori} without writing down the equations). The region of closed trajectories has an infinite volume, which presents difficulties in, e.g., calculation of the effective viscosity of a dilute hard-sphere suspension at the quadratic order in concentration \citep{batchelor_green_1972b}.

%
\begin{figure}[t]
 \centering
\includegraphics[width=0.6\textwidth]{SeparatrixBG.pdf}
\caption{Phase portrait of the trajectories in the symmetry $xz$-plane, within the BG approximation of $z_0=\infty$. All lengths are scaled with the particle radius $a$.
Throughout the paper the reference sphere is at the origin and the trajectories of the second sphere are shown. Due to the fore-and-aft and top-down symmetries only one quadrant is depicted. The two types of the trajectories -- closed (blue) and open (red), are separated by the separatrix, the open trajectory that asymptotically approaches the $x$-axis \citep{lin_lee_sather_1970,batchelor_green_1972}. For two spheres at the same vertical line, the maximal separation for closed trajectories is of order of $10^{-5}$, see \cite{arp}. As a result, at this resolution, the trajectories are indistinguishable when approaching the $z$-axis. This includes the shown open trajectory that crosses the $z$-axis slightly above the closed trajectories. The time-period of revolution along the shown closed trajectory is more than $700$ (here and thereafter the time units of inverse shear rate $\dot{\gamma}^{-1}$ are used).}
\label{fig:SeparatrixBG}
\end{figure}
%
\begin{figure}[t]
 \centering
\includegraphics[width=0.6\textwidth]{Schematics.pdf}
\caption{The phase portrait in the symmetry $xz$-plane at finite $z_0=5$. The fore-and-aft and top-down symmetries survive the wall perturbation in the leading order. The phase portrait exhibits two critical (equilibrium) points: the saddle (hyperbolic) point $r_s$ and the neutral equilibrium (elliptic) point $r_c$, representing a completely different topology from the BG theory in Fig.~\ref{fig:SeparatrixBG}. As $z_0 \to \infty$, the topology of the phase portrait is preserved, while the critical points are being shifted to infinity. }
\label{fig:Schematics}
\end{figure}

We demonstrate here that when a distant wall is considered, the evolution of $\bm r$ remains autonomous in the leading approximation, $\dot{\bm r}=\bm V(\bm r(t))$. Thus at any finite $z_0\gg a$  we can still examine the phase portrait, which is however qualitatively different from that in Fig.~\ref{fig:SeparatrixBG}. Our calculation is not a perturbation theory of the BG solution as we do not assume $\bm V\approx \bm V^0$, so the disturbed phase space flow $\bm V(\bm r)$ is significantly different from $\bm V^0$. The change in topology occurs because at finite $z_0$ there exist critical points at which $\bm V(\bm r)$ vanishes, see the phase portrait in the symmetry plane in Fig.~\ref{fig:Schematics}. The saddle (hyperbolic) point $r_s$, the closer of the two critical points to the origin, is unstable. The other neutral equilibrium (elliptic) point $r_c$ corresponds to a marginally stable configuration, where the pair flows without changing its inter-particle distance and orientation, see Fig.~\ref{Draw1}(a). Not too large deviations from this state result in the sphere orbiting around this elliptic point. These dancing trajectories would have rather unusual appearance when considered in the laboratory frame: while one sphere travels downstream, the other sphere revolves around a point co-moving in space with the first sphere, see Fig.~\ref{Draw1}(b). The phase plane at $x>0$ is characterized by two disconnected regions of closed trajectories shown in Fig.~\ref{fig:Schematics} by the blue curves. The region to the left of
the separatrix (blue) that crosses $x=r_s$ resembles the BG's closed trajectories. The trajectories around the elliptic point $(r_c, 0)$ are solely due to the presence of the wall. The lowest red curve is an open trajectory similar to the BG's: after the encounter the vertical separation is restored to its initial value. In contrast to that, along the (red) trajectories circumventing the elliptic point, the vertical separation of the particles reverses sign after the encounter. We call these ``swapping trajectories", since they seem to correspond to the numerical findings of \cite{zurita-gotor_2007} at $z_0\sim a$, where trajectories with particles swapping their vertical positions after the encounter in a channel flow were reported (careful consideration of the Figure presented in \cite{zurita-gotor_2007} reveals slight changes of the vertical coordinates which seem to be a higher order effect than that considered here). The phase portrait for the evolution of inter-particle distance along the swapping trajectory obtained in \cite{zurita-gotor_2007} numerically, agrees with that predicted here theoretically. To prove that the sign-reversal of the vertical separation predicted here implies swapping, it has to be shown that the center of mass of the pair is not displaced vertically as a result of the encounter. We leave the rigorous proof for future work, focusing here on the evolution of inter-particle distance only. Thus, the use of the term ``swapping trajectories" here, strictly speaking, refers to open trajectories with sign-reversal of the vertical separation following the encounter.

The three-dimensional trajectories are more complex. The circle of radius $r_c$ around the $z$-axis provides the critical curve with $\bm V(\bm r)=0$. The configurations with $\bm r$ on that circle are stationary, so that for instance there is a stationary pair where only the $y$-coordinates of the particles are different. Displacements from these stationary configurations result in closed trajectories that loop around the critical circle, see Fig.~\ref{Fig8b_CircleRc}. In contrast with the symmetry plane, where the BG trajectories display no behavior similar to dancing, some of the three-dimensional BG trajectories do look rather similar (notice that it was not stressed in the original work or in \cite{arp}), see Fig.~\ref{FigX_BG_closed_trajectory1} for comparison.
%
\begin{figure}[t]
 \centering
%\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.45\textwidth]{Draw1.pdf} &
\hskip1cm&
\includegraphics[width=0.45\textwidth]{Draw2.pdf} \\
(a) && (b)
\end{tabular}
%\end{center}
\caption{(a) There is a unique value of the stream-wise separation distance between two particles, $x=r_c$, flowing along the same streamline of the Poiseuille flow, for which they flow steadily without changing their configuration. The stability of such motion is marginal, cf. (b). The value $r_c=4 z_0$ is confirmed in the numerical simulations of the Navier-Stokes equations for $z_0=5$.
(b) Trajectories that pass through points around the stationary point $x=r_c$, $z=0$, crossing $x$-axis at distance larger than $2\sqrt{2}z_0$, exhibit a peculiar dancing dynamics. In the coordinate frame co-moving with the trailing (left) particle, the leading (right) sphere follows an elongated closed orbit around $(r_c, 0)$. Similar trajectories hold outside the symmetry plane.}
\label{Draw1}
\end{figure}

We emphasize the topological difference between the phase portrait in Fig.~\ref{fig:Schematics} and that of the BG theory in Fig.~\ref{fig:SeparatrixBG}. There are two disconnected regions of closed trajectories. In one region the particles orbit each other, similarly to $z_0=\infty$ approximation, however the volume of this region is finite. The other region contains dancing closed trajectories and at large $x^2+y^2$ it is bounded by the surface of revolution $|z|\propto z_0(x^2+y^2)^{-3/2}$. This is similar to the BG bounding surface, $|z|\propto (x^2+y^2)^{-3/2}$, however boosted by the large $z_0$ factor. In both cases the volume of the phase space domain containing closed trajectories diverges, so the divergences in the second order in particle concentration stress calculations of \cite{batchelor_green_1972b} are not regularized by the wall. At finite $z_0$ the two regions of closed trajectories are separated by a region of a new type of open swapping trajectories that, in contrast to the BG theory, cross the $x$-axis. Then the top-down symmetry, which holds remarkably in the presence of the wall, implies that for open trajectories that cross the $x$-axis, the vertical component of the inter-particle distance reverses its sign, as in numerically observed swapping trajectories \cite{zurita-gotor_2007}. At least some features of the presented topology, derived theoretically at $z_0\gg a$, work accurately down to $z_0=5a$, as demonstrated by our in-house numerical simulations of the Navier-Stokes equations.
%
\begin{figure}[t]
 \centering
\includegraphics[width=0.7\textwidth]{Fig8b_CircleRc.pdf}
\caption{The blue line depicts the closed trajectory that forms a loop around the critical circle of radius $r_c$ (whose segment is shown by the green line) for $z_0=5$.
The trajectory can be shrunk to a (necessarily critical) point by continuously changing the initial conditions. The red line shows the BG trajectory
that starts from the same initial condition as the blue line. The period of revolution along the closed trajectory is $1165$.}
\label{Fig8b_CircleRc}
\end{figure}
%
\begin{figure}[t]
 \centering
\includegraphics[width=0.75\textwidth]{FigX_BG_closed_trajectory1.pdf}
\caption{Three-dimensional BG trajectory (red curve) of the sphere in the frame co-moving with the reference sphere. The closed trajectory has a geometrical center on the $y$-axis. In contrast to the trajectory in Fig.~\ref{Fig8b_CircleRc}, this curve cannot be shrunk to a point by a continuous change of the initial conditions. That point would have to be critical and the BG phase space does not admit those. For this trajectory, a distant wall is only a regular small perturbation. % (the trajectories for $z_0=20$ and $z_0=\infty$ are indistinguishable at the resolution used here).
The period of revolution is $120$.}
\label{FigX_BG_closed_trajectory1}
\end{figure}

In the next Section we present the detailed derivation of the evolution equation for the inter-particle distance in the wall-bounded shear flow. Sec.~\ref{singular} demonstrates why the wall presents a singular perturbation of the BG theory. In Sec.~\ref{infinitely} we review the BG trajectories that serve as the reference point of our study. We present the results of the numerical solutions of the derived evolution equation in Sec.~\ref{finitely}. Sec. \ref{dancing-swapping} presents full solution of equation of motion in the dancing-swapping region. Section \ref{dns} presents the confirmation of the theory by direct numerical simulations of the motion of a pair of spheres in the Poiseuille flow. In the last Section we conclude our results, discuss the applicability of the BG theory and formulate some open questions.

\section{Evolution of the distance between two particles transported by a shear flow near wall} \label{interpair}

In this Section we derive the autonomous evolution equation for the distance $\bm r(t)$ between two spheres transported by the Poiseuille flow. We make the simplifying assumption that both spheres are much closer to one of the bounding walls than the other. Thus the particles are effectively transported by the shear flow and not the parabolic velocity profile. The hydrodynamic interaction of particles transported by an unbounded shear flow are well-studied and their velocities
$\bm V^0_{\alpha},\ \alpha=1,2$ were considered in \cite{batchelor_green_1972}. This analysis serves as a starting point of our study. We also assume that the distance to the wall is much larger than the particles' radii. We derive the particles' relative velocity as a sum of $\bm V^0_{\alpha}$ and the correction velocity $\delta \bm V_{\alpha}$. The correction velocity is not necessarily smaller than $\bm V^0$, as our solution is not a perturbation around the solution for an unbounded shear flow.

\subsection{Direct approach} \label{direct}

We set the problem and consider its formulation using the flow for infinitely separated walls as a reference.
The problem of two spheres driven by the Poiseuille flow is described by,
%\begin{eqnarray}&&\!\!\!\!\!\!\!\!
\begin{equation}
  \begin{aligned}
\nabla p\!=\!\eta\nabla^2\bm u,\ \ \nabla \cdot \bm u\!=\!0, \ \ \  \bm u(z\!=\!0)\!=\!\bm u(z\!=\!h)\!=\!0,\ \ \\  %\nonumber\\&&\!\!\!\!\!\!\!\!
u_x(\infty)\!=\!\frac{z(z\!-\!h)\nabla_x p^0}{2\eta},\ \ \bm u(S_{\alpha})\!=\!\bm V_{\alpha}\!+\!\bm \Omega_{\alpha}\!\times\! (\bm x\!-\!\bm x_{\alpha}),
\end{aligned}
\label{ds}
\end{equation}
%\end{eqnarray}
where $\alpha=1, 2$ are the indices of the spheres, $x_{\alpha}$ are the coordinates of the centers and $S_{\alpha}$ is the surface of the $\alpha-$th sphere. We designate the flow by $\bm u$, and the translational and rotational velocities of the spheres by $\bm V_{\alpha}$ and $\bm \Omega_{\alpha}$. The constant pressure gradient $\nabla p^0=-|\nabla_x p^0| \bm{\hat x}$ drives the flow in the positive $x-$direction, $\eta$ is the fluid viscosity, $z$ is the vertical coordinate and $h$ the channel height. We assume that the spheres have equal radii $a$, although most of the calculations below can be done without this assumption.
We will use below $a$ as the unit of length so that the radii are 1.
We assume that the particle inertia is negligible so that the values of $\bm V_{\alpha}$ and $\bm \Omega_{\alpha}$ are determined from the conditions that the total force and torque from the fluid on either particle is zero,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!
\int_{S_{\alpha}} \bm t  dS=0,\ \ \int_{S_{\alpha}} (\bm x-\bm x_{\alpha})\times \bm t  dS=0, \label{inertless}
\end{eqnarray}
where we have introduced the surface traction $\bm t$, which can be written via the stress tensor $\sigma_{ik}$ as,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!
t_i(\bm x)\!=\!\frac{\sigma_{ik}(\bm x\!-\!\bm x_{\alpha})_k}{a},\ \ \sigma_{ik}\!=\!-p\delta_{ik}\!+\!\eta(\nabla_iu_k\!+\!\nabla_ku_i),
\end{eqnarray}
where $\bm x$ belongs to $S_{\alpha}$. We observe that if the spheres are much closer to the wall at $z=0$ than at $z=h$, then we can use different boundary conditions in Eq.~(\ref{ds}),
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!
\bm u(z=0)=0,\ \ \bm u(\infty)\!=\!\dot{\gamma}z\bm{\hat x}, \ \ \dot{\gamma}\!=\!\frac{h|\nabla_x p^0| }{2\eta},
\end{eqnarray}
where we introduced the effective shear rate $\dot{\gamma}$ in terms of the parameters defining the Poiseuille flow. The boundary conditions at $S_{\alpha}$ are unchanged. Without the boundary condition at $z=0$ we reduce to the problem of motion of two spheres in an unbounded shear flow considered in \cite{batchelor_green_1972}. We designate all quantities
of this problem by the superscript zero. Thus $\bm u^0$ is the unbounded shear flow for the two spheres with translational and rotational velocities $\bm V^0_{\alpha}$ and $\bm \Omega^0_{\alpha}$, which obey Eq.~(\ref{inertless}) with $\bm t=\bm t^0$. We look for the solution as superposition of $\bm u^0$ and the flow perturbation $\delta \bm u$ (where the use of $\delta$ does not imply smallness of $\delta \bm u$). We thus have,
%\begin{eqnarray}&&\!\!\!\!\!\!\!\!
\begin{equation}
  \begin{aligned}
\nabla \delta p\!=\!\eta\nabla^2\delta\bm u,\ \ \nabla\cdot\delta\bm u\!=\!0,\ \  \delta\bm u(z\!=\!0)\!=\!-\bm u^0(z\!=\!0),\\ %\nonumber\\&&\!\!\!\!\!\!\!\!
\delta \bm u(\infty)\!=\!0,\ \ \delta \bm u(S_{\alpha})\!=\!\delta \bm V_{\alpha}\!+\!\delta\bm \Omega_{\alpha}\!\times\! (\bm x\!-\!\bm x_{\alpha}),
\end{aligned}
\label{ds1}
\end{equation}
where we introduced deviations of the velocities and of the surface traction from their values in an infinite domain,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!
\delta\bm V_{\alpha}\!=\!\bm V_{\alpha}\!-\!\bm V^0_{\alpha},\ \ \delta\bm \Omega_{\alpha}\!=\!\bm \Omega_{\alpha}\!-\!\bm \Omega^0_{\alpha},\ \ \delta \bm t\!=\!\bm t\!-\!\bm t^0.
\end{eqnarray}
The deviations of the velocities are fixed by the condition that the deviation from the surface traction obeys Eqs.~(\ref{inertless}) with $\delta \bm t$ instead of $\bm t$. We notice that the flow $\bm u^0(z=0)$ in Eq.~(\ref{ds1}) is induced by the spheres, since the unperturbed flow vanishes at $z=0$. Thus $\bm u^0(z=0)$ vanishes at infinity as necessary for consistency of the boundary conditions at the plane and at infinity. For the distant wall the flow $\bm u^0(z\!=\!0)$ can be simplified. This flow obeys the integral representation (see the derivation in Appendix \ref{shear}),
%\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
\begin{equation}
  \begin{aligned}
    u^0_i(\bm x)\!=\!\dot{\gamma}\delta_{i1}z \!-\!\sum_{\alpha}\int_{S_{\alpha}}\!\!\!\frac{Y_{il}(\bm x-\bm x')t^0_{l}(\bm x')dS}{8\pi \eta};\\
    Y_{il}(\bm r)=\frac{\delta_{il}}{r}+\frac{r_ir_l}{r^3},\ \ \bm r=\bm x-\bm x', \label{inter}
    \end{aligned}
\end{equation}
where $Y_{il}$ is the Oseen tensor or the Green's function of the Stokes flow in an unbounded fluid \cite{kim_karrila}. If the vertical positions $z_{\alpha}$ of  the centers of both spheres are much larger than their radii, $a$, then the asymptotic expansion of $\bm u^0(z\!=\!0)$ in $a/z_{\alpha}$ is obtained by Taylor expansion of $Y_{il}(\bm x-\bm x')$ in Eq.~(\ref{inter}) near $\bm x'=\bm x_{\alpha}$. Using the condition of zero force we find that, at the leading order,
%\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
\begin{equation}
  \begin{aligned}
    u^0_i(z=0)\!\approx \!\frac{1}{8\pi \eta}\frac{\partial}{\partial x_m}\sum_{\alpha} Y_{il}(\bm x-\bm x_{\alpha})S^{\alpha}_{lm}|_{z=0};\\
    S^{\alpha}_{lm}\!\equiv \!\!\int_{S_{\alpha}}\!\!\!\!\left(\!(\bm x\!-\!\bm x_{\alpha})_m t^0_l(\bm x)\!-\!\frac{\delta_{ml}(\bm x\!-\!\bm x_{\alpha})_pt^0_p(\bm x)}{3}\!\right)\!dS,
   \end{aligned}
\label{far}
\end{equation}
where the traceless tensor $S^{\alpha}_{lm}$ is \cite{batchelor_green_1972} the force dipole strength of sphere $\alpha$. The $\delta_{lm}$ term  can be added since $\nabla_lY_{il}=0$. The force dipole strengths obey a general form derived in \cite{batchelor_green_1972}. We have $S^{1}_{lm}=S^{2}_{lm}=S_{lm}$ with
%\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!
\begin{equation}
  \small
  \begin{aligned}
\frac{3S_{lm}(\bm r)}{10 \pi \eta a^3\dot{\gamma}}\!=\!\left(\delta_{lx}\delta_{mz}\!+\!\delta_{mx}\delta_{lz}\right)(1\!+\!K)+ \\
%\right.\nonumber\\&&\!\!\!\!\!\!\!\!\!\!\!\left.
 \left(\frac{r_l\left(x\delta_{mz}\!+\!z\delta_{mx}\right)\!+\!r_m\left(x\delta_{lz}\!+\!z\delta_{lx}\right)}{r^2}\! 
-\!\frac{4xz\delta_{lm}}{3r^2}\right)L
%\right.\nonumber\\&&\!\!\!\!\!\!\!\!\!\!\!\left.
\!+\!\frac{2xz}{r^2}\left(\frac{r_lr_m}{r^2}\!-\!\frac{\delta_{lm}}{3}\right) M,\label{dipole}
\end{aligned}
\end{equation}
where the scalar functions $K$, $L$ and $M$ depend on the inter-particle distance $r/a$ only (we omitted the prime in the notation of \cite{batchelor_green_1972}, as the spheres have identical radii in our case). These functions can be completely found only numerically and are considered below as given. We can use Eq.~(\ref{far}) instead of the boundary condition at $z=0$ in Eq.~(\ref{ds1}). The first reflection \cite{hb} gives the leading order approximation for $\delta\bm V_{\alpha}$ as in the Lorentz solution for a sphere in the presence of a distant wall \cite{hb}. The compact expansion can be found below from integral representations.

\subsection{Integral equation for velocities}

Here we derive the integral equation that determines the particle velocities. For future generalization to the case where the distances from the spheres to both walls are comparable we perform the derivation starting from the full formulation given by Eq.~(\ref{ds}).
We use the integral representation of the flow derived in \cite{Fouxon_2017},
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!
u_i(\bm x)\!=\!\frac{\delta_{ix}z(z\!-\!h)\nabla_x p^0}{2\eta}\!-\!\sum_{\alpha}\int_{S_{\alpha}}\!\!\!\frac{S_{il}(\bm x, \bm x')t_{l}(\bm x')dS'}{8\pi \eta},\label{inrepmany}
\end{eqnarray}
where we exploited the symmetry \citep{ps} of Green's function $S_{il}(\bm x, \bm x')=S_{li}(\bm x', \bm x)$. This function is defined by,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
\bm u^S(\bm x)=\frac{1}{8\pi \eta}S_{ik}(\bm x, \bm x_0)g_k, \label{vl}
\end{eqnarray}
where $\bm u^S$ is the Stokeslet flow caused by a point force acting between two parallel plates,
%\begin{eqnarray}&&
\begin{equation}
  \begin{aligned}
-\nabla p^S+\eta \nabla^2 \bm u^S+\bm g\delta(\bm x-\bm x_0)=0,\ \ \nabla\cdot\bm u^S=0, \\  \bm u^S(z=0)=\bm u^S(z=h)=0,\ \ \bm u^S(x^2+y^2\to\infty)=0.\label{stokes1}
  \end{aligned}
\end{equation}
The function $S_{ik}$ is independent of $\bm g$ and it was derived in \cite{LironMochon}. We study the velocities $\bm V_{\alpha}$ using the integral equation for the surface traction $\bm t(\bm x)$ obtained by taking
$\bm x$ in Eq.~(\ref{inrepmany}) to the surface of one of the spheres; this gives
\begin{equation}%&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
  \small
(V_{\alpha})_i+(\bm \Omega_{\alpha}\times (\bm x-\bm x_{\alpha}))_i\!=\!\frac{\delta_{ix}z(z\!-\!h)\nabla_x p^0}{2\eta}
-\!\sum_{\alpha'}\int_{S_{\alpha'}}\!\!\!\frac{S_{il}(\bm x, \bm x')t_{l}(\bm x')dS'}{8\pi \eta}, \label{intr1}
\end{equation}
cf.\ \cite{ps}. This equation holds for all $\bm x$ on $S_{\alpha}$ with $\alpha=1, 2$. Together with the conditions of zero forces and torques it determines $\bm V_{\alpha}$, $\bm \Omega_{\alpha}$ and the surface traction
uniquely \citep{kim_karrila}. We use the assumption $h\gg z_{\alpha}$, meaning that the wall at $z=h$ is much further from the spheres than the one at $z=0$. We can therefore approximately assume
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
S_{il}(\bm x, \bm x')\approx G_{il}(\bm x, \bm x'), \label{condi}
\end{eqnarray}
where $G_{il}(\bm x, \bm x')$ is the Stokeslet near a plane wall defined by
%\begin{eqnarray}&&
\begin{equation} \begin{aligned}
-\nabla p'+\eta \nabla^2 \bm u'+\bm g\delta(\bm x-\bm x')=0,\ \ \nabla\cdot\bm u'=0, \\ \bm u'(z=0)=\bm u'(x\to\infty)=0,\ \
\bm u'(\bm x)=\frac{G_{il}(\bm x, \bm x')g_l}{8\pi \eta}.\label{stokes}
\end{aligned} \end{equation}
The requirement that Eq.~(\ref{condi}) holds when both $\bm x$ and $\bm x'$ belong to the spheres quantifies the assumption that one of the walls is much further than the other. In practice the difference between distances to the upper and lower walls does not have to be too large for the equation to hold. With $z_{\alpha}\ll h$ and this assumption, Eq.~(\ref{intr1}) becomes
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
(V_{\alpha})_i+(\bm \Omega_{\alpha}\times (\bm x-\bm x_{\alpha}))_i\!=\!\dot{\gamma}\delta_{ix}z
-\!\sum_{\alpha'}\int_{S_{\alpha'}}\!\!\!\frac{G_{il}(\bm x, \bm x')t_{l}(\bm x')dS'}{8\pi \eta}. \label{intr}
\end{eqnarray}
We next introduce the decomposition of $G_{il}(\bm x, \bm x')$ into the Stokeslet in an infinite space and the correction due to the wall ${\tilde G}_{il}$,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!
G_{il}(\bm x, \bm x')=Y_{il}(\bm r)+{\tilde G}_{il}(\bm x, \bm x'), \label{do}
\end{eqnarray}
with $\bm r=\bm x-\bm x'$, as above. The contribution ${\tilde G}_{il}$ is induced by the images located at the reflection $(\bm x')^*=(x', y', -z')$ of the source position $\bm x'=(x', y', z')$ with respect to the plane $z=0$. It was found in \cite{blake_1971} that the image singularities are a point force of the same magnitude as the source, but with an opposite sign, a stokes-doublet and a source-doublet, see definitions in the paper. We can write the formula in \cite{blake_1971} as follows
\begin{equation} \begin{aligned}%\begin{eqnarray}&&\!\!\!\!
{\tilde G}_{il}(\bm x, \bm x')=-Y_{il}(\bm R)+2z'G^{(1)}_{il}(\bm R)+2z'^2 G^{(2)_{il}}(\bm R),\\ G^{(1)}_{il}\!=\!\left(2\delta_{3l}\!-\!1\right)\partial_l Y_{i3},\ \
G^{(2)}_{il}\!=\!\frac{\left(1\!-\!2\delta_{3l}\right)\left(R^2\delta_{il}\!-\!3R_iR_l\right)}{R^5},\label{gr}
\end{aligned}\end{equation}
where $\bm R=\bm x-(\bm x')^*$ is the distance from the images and there is no summation over repeated indices. The symmetries of the Green's functions $G_{il}(\bm x, \bm x')=G_{li}(\bm x', \bm x)$ and $Y_{il}(\bm x, \bm x')=Y_{li}(\bm x', \bm x)$ imply the symmetry ${\tilde G}_{il}(\bm x, \bm x')={\tilde G}_{il}(\bm x, \bm x')$, which can be confirmed directly.
